---
title: "Tsibble Exercise"
author: "Maren Rieker & Kabir Sandrolini"
date: "11/1/2021"
output: html_document
---

## Packages Used

```{r}
knitr::opts_chunk$set(echo = TRUE)

library(tsibbledata)
library(tidyverse)
library(tsibble)
library(lubridate)
library(ggplot2)
```

## Retrieving Data

In this example, we use a data set from the tsibbledata package, nyc_bikes
(<https://github.com/tidyverts/tsibbledata>). nyc_bikes contains information
on trips taken in 2018 using NYC rental bikes. The bike_id helps to uniquely 
identify each observation, and the start_time serves as the index. It is an 
**irregular** tsibble, because bike rentals occur at different time intervals.

As a first step, please assign the data frame to an object and check its format.

```{r}
example <- nyc_bikes
example
class(example)

```

## Tidy Data

Tip: For tidying our data, think of the tidy data process Kabir told you about
earlier.

```{r}
#check for duplicates
duplicates(example, index = start_time)

#check for time gaps
has_gaps(example, .full = FALSE)

count_gaps(example, )

example %>% scan_gaps()

example %>% fill_gaps()

#as you may notice, tsibble is quite flexible with the arguments you can
#provide while checking your data for any problems
```

## Working with our Tsibble

#Exercise 1
At the end of this exercise, we want to illustrate for how long different
customers have used the bikes. We want to differentiate customers
by birth year and gender.

```{r}
#clean up the gender column
example %>% count(gender == "Unknown")

bikes_nyc <- filter(example, gender != "Unknown")

bikes_nyc %>% count(gender)

#calculate the renting duration and update the tsibble
class(example$start_time)
class(example$stop_time)

bikes <- bikes_nyc %>%
  group_by_key(c(gender, type)) %>%
  mutate(duration = difftime(stop_time, start_time,
                             units = "mins")) %>%
  update_tsibble(index = duration, 
                 validate = FALSE)
```

## Plot the results for a visual overview of the data

```{r}
class(bikes$duration)
bikes$duration <- as.numeric(bikes$duration)

ggplot(bikes, aes(x = duration)) + geom_histogram() + xlim(0, 150)

ggplot(bikes_nyc, aes(x = birth_year, y = bikes$duration, color = gender)) + 
  geom_point(alpha = 0.2) + scale_x_continuous() + scale_y_time() + 
  ylim(0, 60) + theme(axis.text.x = element_text(angle = 90)) + 
  facet_wrap(~month(bikes_nyc$start_time)) + 
  labs(x = "Year of Birth", y = "Duration of Trip in Minutes", color = 
         "Gender")
```

#Exercise 2
The goal of this exercise is again a visualisation of certain information we can
obtain through the nyc_bikes data set.

Here, you can practice how to set up a tsibble. Because we use the same data
set as before, you do not need to check and tidy the data again. 

```{r}
nycbikes_ts <- nyc_bikes %>% 
   as_tsibble(key = bike_id, index = start_time, regular = FALSE)

```

But we want to transform the temporal data a bit differently this time around - 
measuring the duration of trips in minutes without decimals and compute a 
variable that contains the start hour for each rented bike.

```{r}
nyc2 <- nycbikes_ts %>% mutate(duration_mins = as.numeric(
   round(difftime(stop_time, start_time, units = "mins"),0)))

#index by start_hour
nyc2 %>% 
   index_by(start_hour = floor_date(start_time, unit = "1 hour"))

#extract general start hour and date from more specific timestamp data
nyc3 <- nyc2 %>% 
   mutate(start_hour = floor_date(start_time, unit = "1 hour")) %>% 
   mutate(date = as_date(start_hour), time = hour(start_hour)) %>% 
   filter(gender != "Unknown")

#differentiate again by gender
nyc3 %>% ggplot(aes(x = date, y = duration_mins, colour = gender)) + 
   geom_point(size = 0.5, alpha = 0.5) + ylim(0, 120) + labs(x = "Date", 
              y = "Duration in Minutes", color = "Gender", title = 
                "Length of trips over the year by gender")
```

Now we plot again. This time, please visualise the duration of rentals
noting the different kind of users, customers who have a subscription and 
one-time users.

```{r}
nyc3 %>% ggplot(aes(x = date, y = duration_mins, colour = type)) + 
   geom_point(size = 0.5, alpha = 0.5) + ylim(0, 120) + labs(x = "Date", 
              y = "Duration in Minutes", color = "Type", title = 
                "Length of trips over the year by subscription type")
```

Also very interesting (and great to do with tsibble :-)) is to look at the
rate of people renting bikes over the year.

```{r}
#summarize total number of trips by hour, using the start hour as index
hourly_trips <- nyc2 %>% 
   index_by(start_hour = floor_date(start_time, unit = "1 hour")) %>% 
   summarise(ntrips = n())

#fill any gaps, and add column for date
full_trips <- hourly_trips %>% 
   fill_gaps(ntrips = 0L) %>%
   mutate(date = as_date(start_hour), time = hour(start_hour), 
          month = month(date))

#visualise trips per hour by date over the course of the year
full_trips %>% ggplot(aes(x = date, y = ntrips)) + geom_line() + 
  labs(x = "Date", y = "Trips per hour and bike")

#plotting a smoothing line, we find that the number of trips are highly 
#correlated with the seasons, declining as the weather gets colder.
full_trips %>% ggplot(aes(x = month, y = ntrips)) + geom_smooth() + 
   scale_x_discrete(limits = month.abb) + labs(x = "Month")
```
